---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Sceptic priors and climate consensus

<!-- badges: start -->
<!-- badges: end -->

This repository contains code and data for my working paper, ["Sceptic priors and climate consensus"](https://grantmcdermott.com/papers/scepticpriors.pdf). 

> **Abstract:** How much evidence would it take to convince sceptics that they are wrong about climate change? I explore this question within a Bayesian framework. I consider a group of stylised sceptics and examine how these individuals update their beliefs in the face of current and continuing climate change. I find that available evidence in the form of instrumental climate data tends to overwhelm all but the most extreme priors. Most sceptics form updated beliefs about climate sensitivity that correspond closely to estimates from the scientific literature. However, belief convergence is a non-linear function of prior strength. It becomes increasingly difficult to convince the remaining pool of sceptics that they are wrong. I discuss the necessary conditions for consensus formation under Bayesian learning and show how apparent deviations from the Bayesian ideal still be accommodated within the same conceptual framework. I argue that a generalized Bayesian model thus provides a bridge between competing theories of climate scepticism as a social phenomenon.

Click on the green "Code" button above to clone or download the repo to your local computer. Alternately, click on the "fork" button at the very top right of the page to create an independent copy of the repo within your own GitHub account.

## Reproducibility

I use [**Make**](https://www.gnu.org/software/make/) to automate the entire project. Assuming that you have installed Make and have met all of the other dependencies &mdash; see [below](#dependencies) &mdash; the **TL;DR** version for reproducing everything is:

```sh
## Run these commands in the shell
git clone git@github.com:grantmcdermott/sceptic-priors.git
cd sceptic-priors
make
```

You can also limit the build to subsections of the project by passing Make a relevant meta target, e.g.

- `make data` will construct the dataset
- `make main` will run the main analysis
- `make sensitivity` will run all of the sensitivity analyses etc.

See the [Makefile](Makefile) to get a sense of the options. The associated DAG is [here](makefile-dag.png) (warning: it's complicated).

## Dependencies

While the entire project can be reproduced via a single call to `make`, users must first satisfy various software dependencies. There are two options: i) Manual configuration, or ii) Docker.

### Manual configuration

#### Step 1. Install R and R libraries

With the exception of a tiny bit of Python code, all of the analysis is conducted in the R programming environment. R is free, open-source and available for download [**here**](https://www.r-project.org/). The code has been tested against R version 4.0.2.

Once R is successfully set up on your system, you will need to install a number of external R libraries. I have used [**renv**](https://rstudio.github.io/renv/) to snapshot the project's R environment. To install all of the necessary R libraries, you need simply run the following command from your R console:

```r
## Run these commands in R
# renv::init()    ## Only necessary if you didn't clone/open the repo as an RStudio project
renv::restore()  ## Enter "y" when prompted
```

#### Step 2. Install CmdStan

The workhorse Bayesian regressions for this project are passed from R to [**CmdStan**](https://mc-stan.org/users/interfaces/cmdstan). This enables the Bayesian MCMC computation to complete much, much faster than it would otherwise. While R and CmdStan are two separate programs, the easiest way to install the latter is from the former. Assuming that you have completed Step 1 above, run the following line from your R console:

```r
## Run this command in R
cmdstanr::install_cmdstan(cores = 2)
```

#### Step 3. Optional(ish)

- As mentioned, I use a tiny bit of Python code to extract some [IDL 'save' data](https://pypi.org/project/IDLSave/) as part of the data prep process. Assuming that you have cloned my repo as-is and did not delete any of the data files, Make will automatically skip this section and the Python requirement will be moot. Failing that, however, you will need **SciPy**'s [File IO](https://docs.scipy.org/doc/scipy/reference/tutorial/io.html) module. Regular Python users will almost certainly have SciPy installed on their system already. If not, you can install it yourself, e.g. with PyPi or Conda. I strongly recommend that R users go through **reticulate** (see [here](https://rstudio.github.io/reticulate/articles/python_packages.html)). FWIW, I am using Python 3.8.5 and SciPy 1.5.1 at the time of writing.

- The `extrafont` package is used to embed [Fira Sans](https://fonts.google.com/specimen/Fira+Sans) fonts in the figures. Please note that the Fira Sans font family must be installed separately on your system and also requires some minor setup before R recognizes it (instructions [here](https://github.com/wch/extrafont/blob/master/README.md)). However, you can also skip this setup if you want; R's default Arial fonts will be used if Fira Sans is unavailable.


### Docker

For those of you who don't feel like configuring a manual setup, I also provide a [Docker image](https://hub.docker.com/repository/docker/grantmcd/sceptic) that bundles all of the necessary elements. You'll also want to mount this repo on the running container so that you can access the actual scripts. For example, say that you've cloned this repo to `/home/yourname/sceptic-priors` on your computer. Then you would run

```sh
docker run -it --rm -v /home/yourname/sceptic-priors:/mnt/sceptic-priors grantmcd/sceptic /bin/bash
```

(This will take a few minutes to download the container image from DockerHub the first time you run it. But the container will be ready and waiting for immediate deployment on your system thereafter.)

Once the container is up and running, you should see something like:

```sh
root@53f880d5ea69:/payload# 
```

You can then navigate to the mounted volume and start running the scripts, e.g.

```
root@53f880d5ea69:/payload# cd /mnt/sceptic-priors/ 
root@53f880d5ea69:/mnt/sceptic-priors# Rscript --vanilla R/sceptic.R
```

Of course, you can also open up the interactive R shell (type `R`) instead of running the scripts. Exit and stop the Docker container by typing `exit`.

If you are unfamiliar with Docker, then I recommend starting [here](https://ropenscilabs.github.io/r-docker-tutorial/).


## Performance

```{r perf, include = FALSE}
library(data.table)
library(here)
perf = rbindlist(lapply(list.files(here('performance'), full.names = TRUE), fread))
```


The code has been refactored to run all the Bayesian computation through **CmdStan**. This has yielded *considerable* speed gains, to the point that the entire analysis can be completed in under `r ceiling(sum(perf$sec)/60/10)*10` minutes on my [Dell Precision 5530](https://wiki.archlinux.org/index.php?title=Dell_Precision_5530) laptop.<sup id="a1">[1](#f1)</sup> The table below provides a detailed performance record for the different model runs. Note that I am excluding the data preparation and paper production steps, but each of these only takes a few seconds.

```{r perf_tab, echo = FALSE}
ord = data.table(file = list.files(here('R')))
ord[, run := gsub('.R|-sensitivity', '', file)][, file := paste0('`R/', file, '`')]
setcolorder(ord, c('run', 'file'))
ord[, ord := gsub('[[:alpha:]]|-', '', run)][, run := gsub('^.[0-9]-', '', run)]
ord = ord[!grepl('data', run)]

perf[, sec := as.numeric(sprintf('%.2f', sec))]
perf[run=='evid', run := 'evidence'][run=='rec', run := 'recursive']
perf = ord[perf, on = 'run']
rm(ord)
setorder(perf, ord); perf[, ord:=NULL]
perf[, run2 := fcase(run=='alt-gmst', 'Alternative GMST series',
										 run=='anthro', 'Anthrogenic forcings separate',
										 run=='eff', 'Adjusted forcings efficacies',
										 run=='me-forcings', 'Measurement error in forcings',
										 run=='me-gmst', 'Measurement error in GMST')]
perf[!is.na(run2), run := paste(run2, '(sensitivity)')]
perf[is.na(run2), run := stringr::str_to_title(run)]
perf[, run2 := NULL]
names(perf) = c('Run', 'File', 'Time (sec)', 'Cores used', 'RAM', 'OS', 'Architecture')
knitr::kable(perf)
```


## Problems

If you have any trouble running the code, or find any errors, please file an issue on this repo and I'll look into it.

## License

The software code contained within this repository is made available under the [MIT license](http://opensource.org/licenses/mit-license.php). The data and figures are made available under the [Creative Commons Attribution 4.0 license](https://creativecommons.org/licenses/by/4.0/).

<sup><b id="f1">1</b></sup> Previous versions of the code took a whole day to complete on a cloud server. So you'll understand my being pleased by this improvement. [↩](#a1)