---
title: "Supplementary Material"
subtitle: "Sceptic priors and climate consensus"
author: "Grant McDermott"
# date: "10/14/2020"
bibliography: ../sceptic/sceptic.bib
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    # toc: true
    citation_package: natbib
    dev: cairo_pdf
header-includes: 
  - \usepackage{fontspec}
  - \usepackage{booktabs}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{dcolumn}
  - \newcolumntype{A}{D{.}{.}{2.3}}
  - \usepackage{sectsty} % Allows your to change titles style
  - \allsectionsfont{\sffamily} % Define the style of all titles
mainfont: texgyrepagella-regular.otf
sansfont: Fira Sans
---

\newcommand{\beginsupplement}{%
	\setcounter{table}{0}
	\renewcommand{\thetable}{SM\arabic{table}}%
	\setcounter{figure}{0}
	\renewcommand{\thefigure}{SM\arabic{figure}}%
}
\beginsupplement

\listoftables
\listoffigures
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	cache = TRUE, 
	echo = FALSE, 
	warning = FALSE,
	message = FALSE,
	fig.path = here::here('figs/'),
	fig.align = 'center',
	out.width = '100%',
	dpi=300
	)
```

```{r libs, cache=FALSE, message=FALSE}
library(fst)
library(data.table)
library(ggplot2)
library(ggridges)
library(forcats)
library(hrbrthemes)
library(extrafont)
library(kableExtra)
library(here)

# Functions and global elements ----

## Match short prior names to long prior names
match_priors = function(x) {
	x = gsub('ni', 'Noninformative', x)
	x = gsub('lukemod', 'Moderate Lukewarmer', x)
	x = gsub('lukestrong', 'Strong Lukewarmer', x)
	x = gsub('denmod', 'Moderate Denier', x)
	x = gsub('denstrong', 'Strong Denier', x)
	return(x)
}

## Match short RCP names to long RCP names
match_rcps = function(x) {
  x <- gsub('rcp26', rcp_names[1], x)
  x <- gsub('rcp45', rcp_names[2], x)
  x <- gsub('rcp60', rcp_names[3], x)
  x <- gsub('rcp85', rcp_names[4], x)
  return(x)
}

## Assign colours and names for later graphs ##
rcp_names = c('(a) RCP 2.6', '(b) RCP 4.5', '(c) RCP 6.0', '(d) RCP 8.5')
rcp_cols = scales::viridis_pal(option='plasma')(9)[c(1,3,5,7)] 

prior_names = c('Strong Denier', 'Moderate Denier', 
                'Strong Lukewarmer', 'Moderate Lukewarmer', 'Noninformative')
# c(brewer.pal(12, 'Paired')[c(2, 1, 6, 5)], '#000000') ## Want slightly darker for light pairs
prior_cols = c('Strong Denier'='#1F78B4', 'Moderate Denier'='#8BBDDA',
               'Strong Lukewarmer'='#E31A1C', 'Moderate Lukewarmer'='#F68080',
               'Noninformative'='#000000')

# Plot theme and font ----

## Fira Sans font for figures. Download here: https://bboxtype.com/typefaces/FiraSans/#!layout=specimen
## Must then register with R. See here: https://github.com/wch/extrafont 
## Will revert to ggplot2 default if not installed.
font_type = choose_font('Fira Sans')

## Set global plot theme 
theme_set(
  theme_ipsum(
    # base_size = 12,
    # axis_title_size = 14,
  	axis_title_size = 12,
    axis_title_just = 'c'
    ) +
    theme(
      text = element_text(family = font_type),
      strip.text = element_text(hjust = 0.5)
      )
  )
```

```{r data}
# climate = fread(here('data/climate.csv'))
priors_dt = fread(here('data/priors.csv'))
# params = fread(here('results/main/params.csv'))
# tcr = read_fst(here('results/main/tcr.fst')); setDT(tcr)
# gmst_pred = fread(here('results/main/gmst-pred.csv'))
# gmst2100 = read_fst(here('results/main/gmst2100.fst')); setDT(gmst2100)
# tcr_rec = fread(here('results/recursive/tcr-rec.csv'))
# evid = fread(here('results/evidence/evid.csv'))
scc = fread(here('results/PAGE09/scc.csv'))
```


# Tables

## Sensitivity analysis

\begin{table}[ht] \centering 
	\caption{TCR efficacies used in ``MEA'' I and II sensitivity runs} 
	\label{tab:marvel} 
	\begin{threeparttable} 
		\begin{tabularx}{.75\textwidth}{@{\extracolsep{5pt}} Xcc}
			\toprule
			Forcing agent & Mean &   95\% C.I.   \\ 
			\midrule
			Aerosols      & 1.55 & (1.05, 2.05)  \\
			GHGs          & 1.17 & (1.07, 1.28)  \\
			Land use      & 3.82 & (-2.16, 9.80) \\
			Ozone         & 0.66 & (0.34, 0.98)  \\
			Solar         & 1.68 & (-1.27, 4.63)  \\
			Volcanic      & 0.61 & (0.33, 0.89)  \\ 
			\bottomrule
		\end{tabularx} 
		\begin{tablenotes}
			\footnotesize
			\item Notes: Adapted from Table S1 of \cite{marvel2016implications}. Confidence intervals on the sample means are constructed from a \textit{t} distribution with 4 degrees of freedom.
		\end{tablenotes}
	\end{threeparttable} 
\end{table} 

## Future temperatures

\begin{table}[h] \centering 
	\caption{Covariate vectors for 2100 predictions} 
	\begin{threeparttable} %%% added %%% 
		\begin{tabularx}{.75\textwidth}{@{\extracolsep{1pt}} X A A A A } 
			%\begin{tabularx}{\textwidth}{X c c c c} 
			%\hline\hline
			\toprule
			&\multicolumn{1}{c}{RCP 2.6}&\multicolumn{1}{c}{RCP 4.5}&\multicolumn{1}{c}{RCP 6.0}&\multicolumn{1}{c}{RCP 8.5}\\
			%&\multicolumn{1}{c}{\footnotesize420 ppmv CO$_2$}&\multicolumn{1}{c}{\footnotesize540 ppmv CO$_2$}&\multicolumn{1}{c}{\footnotesize670 ppmv CO$_2$}&\multicolumn{1}{c}{\footnotesize940 ppmv CO$_2$}\\
			%\hline
			%\\[-1.8ex] 
			\midrule
			$RF_{2100}$  																			& 2.626		& 4.281		& 5.522		& 8.340		\\ 
			\hspace{5 pt} \textit{CO$_2$ component}	& \multicolumn{1}{c}{\textit{\hspace{1em}85\%}}	&	\multicolumn{1}{c}{\textit{\hspace{1em}83\%}}	&	\multicolumn{1}{c}{\textit{\hspace{1em}86\%}}	&	\multicolumn{1}{c}{\textit{\hspace{1em}78\%}}	\\
			\hspace{5 pt} \textit{Solar component}		& \multicolumn{1}{c}{\textit{\hspace{1em} 7\%}}	&	\multicolumn{1}{c}{\textit{\hspace{1em} 4\%}}	&	\multicolumn{1}{c}{\textit{\hspace{1em} 3\%}}	&	\multicolumn{1}{c}{\textit{\hspace{1em} 2\%}}	\\
			\\[-1.8ex] 
			$\overline{VOLC}$																	& 0.017		& 0.017		& 0.017		& 0.017		\\
			\\[-1.8ex] 
			$\overline{SOI}$																	&\text{-}0.079	&\text{-}0.079	&\text{-}0.079	&\text{-}0.079	\\
			\\[-1.8ex] 
			$\overline{AMO}$ 																	&\text{-}0.002 &\text{-}0.002	&\text{-}0.002	&\text{-}0.002	\\
			%\hline\hline
			\bottomrule
		\end{tabularx}
		\begin{tablenotes}
			\footnotesize
			\item \textit{Notes:} Covariates are used to predict the global mean surface temperature anomaly in the year 2100. The Representative Concentration Pathways (RCPs) are a family of forcing scenarios developed for the IPCC \cite{van2011rcp}. Each RCP has a core component of atmospheric CO$_2$ concentrations, measured in parts per million volume (ppmv). With regard to the covariates in the regression model, total radiative forcing ($RF$) and volcanic aerosols ($VOLC$) are measured in Wm$^{-2}$. The Southern Oscillation Index ($SOI$) and Atlantic Multidecadal Oscilliation ($AMO$) are measured as scaled indices. Future values for $RF$ are taken from the RCP database. For the rest, historical mean values are used.
		\end{tablenotes}
	\end{threeparttable} 
	\label{tab:covariate}
\end{table}


\newpage
\pagebreak

# Figures

## Sensitivity analysis

Figs. \ref{fig:sens_cw14} -- \ref{fig:sens_co2} provide additional context and information related to the various sensitivity analyses undertaken in Section 5.2 of the main text. In each case, the figure caption references against the key listed in first column of Table 4. The figures themselves are directly comparable with Fig. 1 and the same general notes apply (dashed lines denote TCR priors, solid lines denote TCR posteriors, etc.) Note that in some cases the x-axis has been truncated to preserve this direct comparability, even though the posterior distributions may extend beyond the -1 °C to 3 °C range.

```{r tcr_plot}
tcr_plot =
  function(tcr_dt, p_dt = priors_dt, priors_only = FALSE) {
    
    ## Can't use stat_function() that maps to facets, ridges or other aesthetic 
    ## elements. So we have to create the data manually instead. See:
    ## https://github.com/tidyverse/ggplot2/issues/2357
    
  	## Priors
  	p_dt[, prior := factor(match_priors(paste0(prior_type, convic_type)), 
  												 levels=prior_names)]
    p_dt[prior_type=='ni', ':=' (mu = mu*3.71, sigma = sigma*3.71)]
    prior_dens = 
    	rbindlist(lapply(
    		prior_names, 
    		function(x){
    			dt = p_dt[prior==x]
    			tcr_grid = seq(from = qnorm(0.0001, dt$mu, dt$sigma), 
    										 to = qnorm(0.9999, dt$mu, dt$sigma), 
    										 length=100)
    			data.table(
    				tcr = tcr_grid,
    				height = dnorm(tcr_grid, mean = dt$mu, sd = dt$sigma),
    				prior = x
    				)
    			}))
    
    ## Posteriors
    tcr_dens = tcr_dt[, .(data = list(.SD)), by = prior]
    tcr_dens[, dens := lapply(data, function(d) density(d$tcr))]
    tcr_dens[, tcr := lapply(dens, function(d) d$x)]
    tcr_dens[, height := lapply(dens, function(d) d$y)]
    ## Unnest
    tcr_dens = tcr_dens[, .(tcr = tcr[[1]], height = height[[1]]), by=prior]
    tcr_dens[, prior := factor(match_priors(prior), levels=prior_names[c(1,3,2,4,5)])]
    
    ggplot(
    	tcr_dens,
    	aes(x=tcr, y = prior, height=height, group=prior, 
    			col=prior, fill=prior)
    	) +
      ## Dummy data (need to plot first otherwise annotate geom doesn't work)
      geom_density_ridges(stat = 'identity', scale = 2, alpha = 0, col = NA) +
      ## IPCC "likely" region (1.0–2.5 °C)
      annotate('rect', xmin = 1, xmax = 2.5, ymin = 0, ymax = Inf, alpha = .2) +
      ## Priors
      geom_density_ridges(
        stat = 'identity', scale = 1.75, 
        data = prior_dens,
        lty = 2, fill = NA
        ) +
      ## Posteriors
      {if (!priors_only) {
      	geom_density_ridges(stat = 'identity', scale = 1.75, alpha = 0.5, lwd = 0.5)
      	}} +
      ## Stylistic elements
      labs(x = expression('TCR'~'('*degree*C*')'), y = 'Density') +
      xlim(-1, 3) +
      scale_colour_manual(values = prior_cols, aesthetics = c('colour', 'fill')) +
      theme(
        axis.text.y = element_text(vjust = 0),
        axis.title.y = element_blank(),
        legend.position = 'none'
      ) 
    }
```

```{r sens_cw14, dependson = 'tcr_plot', fig.cap = 'TCR densities: "CW14" sensitivity run.'}
alt_gmst = read_fst(here('results/sensitivity/tcr-alt-gmst.fst')); setDT(alt_gmst)
tcr_plot(alt_gmst[series=='cw'])
```

```{r sens_gistemp, dependson = 'tcr_plot', fig.cap = 'TCR densities: "GISTEMP" sensitivity run.'}
tcr_plot(alt_gmst[series=='giss'])
```

```{r sens_hadcrut_me, dependson = 'tcr_plot', fig.cap = 'TCR densities: "HadCRUT ME" sensitivity run.'}
hadcrut_me = read_fst(here('results/sensitivity/tcr-me-gmst.fst')); setDT(hadcrut_me)
tcr_plot(hadcrut_me)
```

```{r sens_df18, dependson = 'tcr_plot', fig.cap = 'TCR densities: "DF18" sensitivity run.'}
df18 = read_fst(here('results/sensitivity/tcr-me-forcings.fst')); setDT(df18)
tcr_plot(df18)
```

```{r sens_mea_i, dependson = 'tcr_plot', fig.cap = 'TCR densities: "MEA I" sensitivity run.'}
mea_i = read_fst(here('results/sensitivity/tcr-eff1.fst')); setDT(mea_i)
tcr_plot(mea_i)
```

```{r sens_mea_ii, dependson = 'tcr_plot', fig.cap = 'TCR densities: "MEA II" sensitivity run.'}
mea_ii = read_fst(here('results/sensitivity/tcr-eff2.fst')); setDT(mea_ii)
tcr_plot(mea_ii)
```

```{r sens_anthro, dependson = 'tcr_plot', fig.cap = 'TCR densities: "Anthro" sensitivity run.'}
anthro = read_fst(here('results/sensitivity/tcr-anthro.fst')); setDT(anthro)
tcr_plot(anthro[series=='anthro'])
```

```{r sens_co2, dependson = 'tcr_plot', fig.cap = 'TCR densities: "CO2" sensitivity run.'}
tcr_plot(anthro[series=='co2'])
```

\pagebreak

## Welfare implications and the social cost of carbon

```{r scc-fig, fig.cap = 'Social cost of carbon (US\\$2005 per ton). The results for each agent type are obtained from 10,000 simulation runs of PAGE \\cite{hope2011page09}. Posterior TCR distributions serve as key inputs to the model, while the remaining parameters are set to the PAGE model defaults. The x-axis is truncated at 100 to aid visual inspection; the uppermost tails of the distributions being well in excess of the range given here.', fig.pos = 'h'}
scc = melt(scc, measure.vars = names(scc), 
					 variable.name = 'prior', value.name = 'scc')
scc[, prior := factor(match_priors(prior), levels=prior_names)]
ggplot(scc, aes(x = scc, col = prior)) +
      geom_line(stat = "density") +
      xlim(-10, 100) + ## NB: x-axis is truncated to aid visual inspection!
      labs(
        x = "US$ (2005)", 
        y = "Density") + 
      scale_colour_manual(values = prior_cols) +
      guides(col = guide_legend(nrow = 2)) +
      theme(
        legend.position = "bottom",
        legend.title = element_blank()
        )
```

\newpage
